Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-09 paper/disc20clustering.tex
Distributed

\section{Discussion and Future Work}
The algorithm PPL from Section~\ref{sec:ppl} can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.
The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
\medskip

In the rest of this section, we propose the distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
The algorithm performs two non-local operations:
\begin{enumerate}
	\item Broadcasting the end of the phase.\label{it:broadcast}
	\item Component merge.
	\label{it:merge}
\end{enumerate}

If a component reaches size $4$ or fresh cluster is not found, we conclude that no perfect partition exists.
Then, we end the phase and broadcast (\maciek{TODO: efficient broadcast cite}) the message to  reset the counters in all clusters.

A merge of two components may require finding a third cluster that contains $3$ alone nodes (for details see the proof of Lemma~\ref{lem:1req})).
We say that a cluster containing $3$ alone nodes is \emph{fresh}.
In the following, we show how to efficiently find a fresh cluster in a distributed manner.
We organize the clusters into an arbitrary rooted balanced binary tree (\maciek{TODO: how to do it efficiently? Does not matter much, as this is performed once at the beginning}), and we broadcast the root to each cluster.
Each cluster maintains the counter of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.

The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log(\ell))$ communication.
Each of these may occur, on average, every $\alpha$ external requests.

>>>>

\section{Discussion and Future Work}

While we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 
The algorithm PPL from Section~\ref{sec:ppl} can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.
The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
\medskip
Now, we propose the distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
The algorithm performs two non-local operations:
\begin{enumerate}
	\item Broadcasting the end of the phase.\label{it:broadcast}
	\item Component merge.
	\label{it:merge}
\end{enumerate}

If a component reaches size $4$ or a fresh cluster is not found, we conclude that no perfect partition exists.
Then, we end the phase and broadcast the message to  reset the counters in all clusters.

A merge of two components may require finding a third cluster that contains $3$ isolated nodes (for details see the proof of Lemma~\ref{lem:1req})).
We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
In the following, we show how to efficiently find a fresh cluster in a distributed manner.
We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
Each cluster maintains the counter of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.

The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log \ell)$ communication.
Each of these may occur, on average, every $\alpha$ external requests.

This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a general partitioning model and for a special learning model.

While are bounds are tight, there are several interesting avenues for future research.
In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
